Categories of Optics

### 2018-05-04 paper.tex
Counit symmetry
Extra tikz code not necessary
Progress

\usepackage{tikz-cd}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{cd}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\tikzset{ed/.style={auto,inner sep=2pt,font=\scriptsize}} %edges
\tikzset{>=stealth'}
\tikzset{vert/.style={draw,circle, minimum size=6mm, inner sep=0pt, fill=white}}
\tikzset{vertbig/.style={draw,circle, minimum size=8mm, inner sep=0pt, fill=white}}
\input{tikzcurverel}

@@

\subsection{Notation}
\todo{I am open to changing any/all of these}
\begin{itemize}
\item Given $f : A \to B$ and $g : A \to C$, I write $[f, g] : A \to B \times C$ and $!_A : A \to 1$.
\item Given $f : A \to B$ and $g : C \to D$, I write $f \times g : A \times C \to B \times D$.
\item Given a $x \in P(C, C)$, I write $\langle x \rangle \in \int^C P(C, C)$ for its image, and $\copr_C : P(C,C) \to \int^C P(C, C)$
\item For a monoidal category, the structure maps are $\alpha$, $\lambda$ and $\rho$. A symmetry is $s$.
\end{itemize}

@@

\todo{ Notation: For a monoidal category, the structure maps are $\alpha$, $\lambda$ and $\rho$, symmetry is $s$. For an action of $(\M, \otimes, I)$ on $\C$ we have two structure maps that are isomorphisms: $\epsilon_A : A \to IA$ and $\mu_{M,N,A} = MNA \to (M \otimes N)A$. }

@@

\begin{remark}
\todo{I hope I can get away with this...} In what follows, we use the standard cheat of omitting associativity isomorphisms.
\end{remark}

@@

If $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, we write $\langle l, r \rangle : (S, S') \hto (A, A')$ for the image of this pair in $\Optic_\C((S, S'), (A, A'))$, and say that $M$ is the complement for this representative. Optics will always be written with a crossed arrow $\hto$, to distinguish them from morphisms of $\C$.

@@

Graphically, this is placing the second optic into the middle of the first:
\begin{center}
\input{diagrams/generic-optic-noline.tikz}
\qquad\raisebox{1.5cm}{$\circ$}\qquad
\input{diagrams/generic-optic2-noline.tikz} \\
\raisebox{1.5cm}{$:=$}\qquad
\input{diagrams/optic-composition.tikz}
\end{center}

The identity optic $(S, S') \hto (S, S')$ is given by $\langle \lambda^{-1}_S, \lambda_{S'} \rangle$, where $\lambda_S : I \otimes S \to S$ is the left unitor for $S$: 
\begin{center}
\input{diagrams/identity-optic-full.tikz}
\end{center}
This dashed line above the diagram represents the unit object. As is standard, we will elide the unit object and unitors in future diagrams. The diagram for the identity becomes simply:
\begin{center}
\input{diagrams/identity-optic.tikz}
\end{center}

@@

Categories of optics admit their own diagrammatic language, by virtue of being \emph{teleological categories}, monoidal categories with counits. In fact, we show in Theorem \ref{optic-is-free-teleological-cat} that $\Optic_\C$ is the free teleological category on a collection of dualisable morphisms $\C$.

In this section we assume that $\C$ is a \emph{strict} symmetric monoidal category. It will be convenient to equip $\Optic_\C$ with a slightly different symmetric monoidal structure: \todo{And the universal property given in this section is an argument for this being the ``morally correct'' tensor}

@@

\begin{proposition}
$\Optic_\C$ equipped with the switched tensor is monoidally equivalent to it equipped with the unswitched tensor.
\end{proposition}
\begin{proof}
\todo{todo, I hope this is true?}
\end{proof}

\begin{definition}[{\cite[Definition 5.1]{CoherenceForLenses}}]
A \emph{teleological category is} a symmetric monoidal category $(\C, \otimes, I)$, equipped with:
\begin{itemize}
\item A wide symmetric monoidal subcategory $\C_d$ of \emph{dualisable morphisms}, with an involutive strong symmetric monoidal functor $(-)^* : \C_d \to \overline{\C_d^\op}$, where---not finding a standard symbol for such a thing---we mean $\overline{\C_d^\op}$ to be the category with both the direction of the arrows \emph{and} the order of the tensor flipped: $(A \otimes B)^* = B^* \otimes A^*$. \todo{Do we want this to be strict?}
\item A symmetric monoidal extranatural family of morphisms $\varepsilon_X : X \otimes X^* \to I$, called \emph{counits}, natural with respect to the morphisms in $\C_d$.
\end{itemize}
\end{definition}
Unpacking the definition, $\varepsilon$ being a symmetric monoidal extranatural transformation amounts to the following diagrams in $\C$ commuting:
\[
\begin{tikzcd}
X \otimes Y^* \ar[r, "f \otimes Y^*"]  \ar[d, "X \otimes f^*", swap] & Y \otimes Y^* \ar[d, "\varepsilon_Y"] \\
X \otimes X^* \ar[r, "\varepsilon_X", swap] & I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X^* \otimes X \ar[r, "\sigma"]  \ar[dr, "\varepsilon_{X^*}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
& I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X \otimes Y \otimes Y^* \otimes X^* \ar[r, "X \otimes \varepsilon_Y \otimes X^*"]  \ar[dr, "\varepsilon_{X \otimes Y}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
& I
\end{tikzcd}
\] 
where $f : X \to Y$ is in $\C_d$.

@@

\begin{proposition}
$\Optic_\C$ forms a teleological category, where:
\begin{itemize}
\item The dualisable morphisms are all morphisms of the form $\iota(f, g)$;
\item The involution is given on objects by $(S, S')^* := (S', S)$, and on a morphisms by \[\iota(f, g)^* := \iota(g, f);\]
\item The counit $\varepsilon_{(S, S')} : (S, S') \switched (S, S')^* = (S \otimes S', S \otimes S') \to (I, I)$ is given by the connector: \[\varepsilon_{(S, S')} := c_{S \otimes S'}.\]
\end{itemize}
\end{proposition}
\begin{proof}
That morphisms of the form $\iota(f, g)$ constitute a category and that $(-)^*$ is a symmetric monoidal involution is clear: we have
\begin{align*}
\left( (S, S') \switched (T, T') \right)^* 
&= \left( S \otimes T, T' \otimes S' \right)^* \\
&= \left(T' \otimes S', S \otimes T  \right) \\
&= (T', T) \otimes (S', S) \\
&= (T, T')^* \otimes (S, S')^*
\end{align*}

To check extranaturality of $\varepsilon$, suppose we have a dualisable optic $\iota(f, g) : (S, S') \hto (T, T')$, so $f : S \to T$ and $g : T' \to S'$. Extranaturality is witnessed by the equality of the string diagrams:
\begin{center}
\input{diagrams/counit-extranatural-left.tikz}
\qquad \raisebox{1.5cm}{$=$} \qquad 
\input{diagrams/counit-extranatural-right.tikz}
\end{center}
Symmetry:
\begin{center}
\includegraphics[width=.4\textwidth]{diagrams/counit-symmetry}
\end{center}
Monoidality:
\begin{center}
\includegraphics[width=.4\textwidth]{diagrams/counit-monoidal}
\end{center}

@@

\begin{proposition}
The functor $\Optic : \SymmMonCat \to \SymmMonCat$ of Proposition \ref{prop-optic-functor} has its image in the subcategory $\Tele$.
\end{proposition}
\begin{proof}
We must show that for a symmetric monoidal functor $F : \C \to \D$, the induced functor $\Optic_\C \to \Optic_\D$ is teleological.

\todo{todo}
\end{proof}

@@

\begin{proposition}
\label{prop-optic-decompose}
Suppose $\langle l, r \rangle : (S, S') \hto (A, A')$ has complement $M$. Then 
\begin{align*}
\langle l, r \rangle = (\varepsilon_{(M, I)} \otimes (A, A'))(j(l) \otimes j(r)^*)
\end{align*}
where $j : \C \to \Optic_\C$ is the functor $j(A) := \iota(A, I)$.
\end{proposition}
\begin{proof}
First note that because $\C$ is strict monoidal, the counit $\varepsilon_{(M, I)} : (M \otimes I, M \otimes I) = (M, M) \hto (I, I)$ is equal to the connector $c_M : (M, M) \hto (I, I)$.
\todo{todo}
\end{proof}

\begin{theorem}
\label{optic-is-free-teleological-cat}
Suppose $\C$ is a strict symmetric monoidal category and $\T$ is a teleological category. For all monoidal functors $F : \C \to \T_d$, there exists a teleological functor $K : \Optic_\C \to \T$, unique up to isomorphism, such that $Kj \cong F$.
%$\Optic$ is left adjoint to $(-)_d$, so $\Optic_\C$ is the free teleological category on a category of dualisable morphisms $\C$.
\end{theorem}
\begin{proof}
On objects, define $K(S, S') = FS \otimes (FS')^*$. Suppose $\langle l, r \rangle : (S, S') \hto (A, A')$ is an optic. By the previous Proposition,
\begin{align*}
\langle l, r \rangle = (\varepsilon_{(M, I)} \otimes (A, A'))(j(l) \otimes j(r)^*)
\end{align*}
But now, in order for $K$ to be a teleological functor with $Kj \cong F$, each piece of this expression must be preserved by $K$, so we are forced to define:
\begin{align*}
K\langle l, r \rangle = (\varepsilon_{FM} \otimes FA \otimes (FA')^*)(Fl \otimes (Fr)^* )
\end{align*}
%
%This is much easier to understand in the diagrammatic language for teleological categories: \todo{todo}
%
%We now have to check that this is well defined. \todo{todo}

\todo{incomplete}
\end{proof}

@@

@@

So in the above example, we would take $F = \mathtt{Int} \times -$ and $G$ to be the identity functor.

@@

@@

\section{Indexed Optics}
\todo{
This possibly looks like
\begin{align*}
\int^{M \in \M} \C(S, M(IA)) \times \C(M A', S')
\end{align*}
with another monoidal action $I$. The $I$ would need to distribute over $M$ in some way. ``Coindexed'' optics seem to have 
\begin{align*}
\int^{M \in \M} \C(S, J(MA)) \times \C(M A', S')
\end{align*}
which is why they pass information `in the other direction' but still transform covariantly.
}

@@

\begin{example}
A minimal example of a prelens that obeys $\fget\fput$ and $\fput\fget$ but not $\fput\fput$ is the following. Let $p : \{A,B,C\} \hto \{X, Y\}$ be the prelens with
\begin{align*}
\fget : \{A,B,C\} &\to \{X, Y\} \\
A &\mapsto X \\
B &\mapsto X \\
C &\mapsto Y 
\end{align*}
\begin{align*}
\fput : \{A,B,C\} \times \{X, Y\} &\to \{A,B,C\} \\
(A,X) &\mapsto A \\
(B,X) &\mapsto B \\
(C,X) &\mapsto B \\
(A,Y) &\mapsto C \\
(B,Y) &\mapsto C \\
(C,Y) &\mapsto C
\end{align*}
\end{example}

>>>>

\usepackage{tikz-cd}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{cd}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.markings}
\tikzset{ed/.style={auto,inner sep=2pt,font=\scriptsize}} %edges
\tikzset{>=stealth'}
\tikzset{vert/.style={draw,circle, minimum size=6mm, inner sep=0pt, fill=white}}
\tikzset{vertbig/.style={draw,circle, minimum size=8mm, inner sep=0pt, fill=white}}
\tikzset{->-/.style={decoration={
  markings,
  mark=at position #1 with {\arrow{>}}},postaction={decorate}}}

@@

@@

@@

\begin{remark}
In what follows, we use the standard cheat of omitting associativity isomorphisms.
\end{remark}

@@

Because this coend takes place in $\Set$, we can describe the set $\Optic_\C((S, S'), (A, A'))$ explicitly. It is the set of pairs $(l, r)$, where $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, quotiented by the relation
\begin{align*}
((f \otimes A) l, r) \sim (l, r (f \otimes A'))
\end{align*}
where $l : S \to M \otimes A$, $r : N \otimes A' \to S'$ and $f : M \to N$.

If $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, we write $\langle l, r \rangle : (S, S') \hto (A, A')$ for the image of this pair in $\Optic_\C((S, S'), (A, A'))$, and say that $M$ is the complement for the representative. Optics will always be written with a crossed arrow $\hto$, to distinguish them from morphisms of $\C$.

@@

Graphically, this is placing the first optic into the hole of the second:
\begin{center}
\input{diagrams/generic-optic-noline.tikz}
\qquad\raisebox{1.5cm}{$\circ$}\qquad
\input{diagrams/generic-optic2-noline.tikz} \\
\raisebox{1.5cm}{$:=$}\qquad
\input{diagrams/optic-composition.tikz}
\end{center}

The identity optic $(S, S') \hto (S, S')$ is given by $\langle \lambda^{-1}_S, \lambda_{S'} \rangle$, where $\lambda_S : I \otimes S \to S$ is the left unitor for $S$: 
\begin{center}
\input{diagrams/identity-optic-full.tikz}
\end{center}
This dashed line above the diagram represents the unit object. As is standard, we will elide the unit object and unitors in future diagrams unless they are necessary. The diagram for the identity becomes simply:
\begin{center}
\input{diagrams/identity-optic.tikz}
\end{center}

@@

In this section we establish a universal property of the $\Optic$ construction. The idea is that every optic $\langle l, r\rangle : (S, S') \hto (A, A')$ consists of a morphism $S \to M \otimes A$ and the `formal dual' of a morphism $M \otimes A' \to S'$, composed with a `formal counit' that traces out the object $M$:
\begin{center}
\input{diagrams/generic-optic-folded.tikz}
\end{center}

We assume temporarily that $\C$ is a \emph{strict} symmetric monoidal category. It will be convenient to equip $\Optic_\C$ with a slightly different symmetric monoidal structure: \todo{And the universal property given in this section is an argument for this being the ``morally correct'' tensor}

@@

\begin{proposition}
$(\Optic_\C, \switched, (I, I))$ is a symmetric monoidal category that is monoidally equivalent to $\Optic_\C$ with the unswitched tensor.
\end{proposition}
\begin{proof}
The proof that $(\Optic_\C, \switched, (I, I))$ is symmetric monoidal is nearly identical to that for the unswitched tensor. Note that due to the switching, the structure morphisms are slightly different:
\begin{align*}
\alpha_{(R, R'), (S, S'), (T, T')} &:= \iota(\alpha_{R,S,T}, \alpha_{T',S',R'}^{-1}) \\
\lambda_{(S, S')} &:= \iota(\lambda_{S}, \rho_{S'}^{-1}) \\
\rho_{(S, S')} &:= \iota(\rho_{S}, \lambda_{S'}^{-1}) \\
s_{(S, S'), (T, T')} &:= \iota(s_{S, T}, s_{S', T'})
\end{align*}

The categories are monoidally equivalent via the identity functor $I : (\Optic_\C, \switched, (I, I)) \to (\Optic_\C, \otimes, (I, I))$, where the structure isomorphisms are given by
\begin{align*}
\iota(\id_{S \otimes S'}, s_{T, T'}) : I(S, S') \otimes I(T, T') \to I((S, S') \switched (T, T'))
\end{align*}

\begin{remark}
Just as in the unswitched case, if $\C$ is a strict monoidal category than so is $(\Optic_\C, \switched, (I, I))$.
\end{remark}

\begin{definition}[Compare {\cite[Definition 5.1]{CoherenceForLenses}}]
A \emph{teleological category is} a symmetric monoidal category $(\C, \otimes, I)$, equipped with:
\begin{itemize}
\item A wide symmetric monoidal subcategory $\C_d$ of \emph{dualisable morphisms}, with an involutive strong symmetric monoidal functor $(-)^* : \C_d \to \overline{\C_d^\op}$, where---not finding a standard symbol for such a thing---we mean $\overline{\C_d^\op}$ to be the category with both the direction of the arrows \emph{and} the order of the tensor flipped: $(A \otimes B)^* = B^* \otimes A^*$. \todo{Do we want this to be strict?}
\item A symmetric monoidal extranatural family of morphisms $\varepsilon_X : X \otimes X^* \to I$, called \emph{counits}, natural with respect to the morphisms in $\C_d$.
\end{itemize}
\end{definition}
Unpacking the definition, $\varepsilon$ being a symmetric monoidal extranatural transformation amounts to the following diagrams in $\C$ commuting:
\[
\begin{tikzcd}
X \otimes Y^* \ar[r, "f \otimes Y^*"]  \ar[d, "X \otimes f^*", swap] & Y \otimes Y^* \ar[d, "\varepsilon_Y"] \\
X \otimes X^* \ar[r, "\varepsilon_X", swap] & I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X^* \otimes X \ar[r, "\sigma"]  \ar[dr, "\varepsilon_{X^*}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
& I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X \otimes Y \otimes Y^* \otimes X^* \ar[r, "X \otimes \varepsilon_Y \otimes X^*"]  \ar[dr, "\varepsilon_{X \otimes Y}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
& I
\end{tikzcd}
\] 
where $f : X \to Y$ is in $\C_d$.

@@

\begin{proposition}
$\Optic_\C$ forms a teleological category, where:
\begin{itemize}
\item The dualisable morphisms are all morphisms of the form $\iota(f, g)$;
\item The involution is given on objects by $(S, S')^* := (S', S)$, and on a morphisms by \[\iota(f, g)^* := \iota(g, f);\]
\item The counit $\varepsilon_{(S, S')} : (S, S') \switched (S, S')^* = (S \otimes S', S \otimes S') \to (I, I)$ is given by the connector: \[\varepsilon_{(S, S')} := c_{S \otimes S'}.\]
\end{itemize}
\end{proposition}
\begin{proof}
That morphisms of the form $\iota(f, g)$ constitute a category and that $(-)^*$ is a symmetric monoidal involution is clear: we have
\begin{align*}
\left( (S, S') \switched (T, T') \right)^* 
&= \left( S \otimes T, T' \otimes S' \right)^* \\
&= \left(T' \otimes S', S \otimes T  \right) \\
&= (T', T) \switched (S', S) \\
&= (T, T')^* \switched (S, S')^*
\end{align*}

To check extranaturality of $\varepsilon$, suppose we have a dualisable optic $\iota(f, g) : (S, S') \hto (T, T')$, so $f : S \to T$ and $g : T' \to S'$. Extranaturality is witnessed by the equality of the string diagrams:
\begin{center}
\input{diagrams/counit-extranatural-left.tikz}
\qquad \raisebox{1.5cm}{$=$} \qquad 
\input{diagrams/counit-extranatural-right.tikz}
\end{center}
Symmetry:
\begin{center}
\input{diagrams/counit-symmetry-left.tikz}
\qquad \raisebox{1.5cm}{$=$} \qquad 
\input{diagrams/counit-symmetry-right.tikz}
\end{center}
Monoidality: \todo{there is basically nothing to do in the graphical calculus}

@@

\begin{proposition}
The functor $\Optic : \SymmMonCat \to \SymmMonCat$ of Proposition \ref{prop-optic-functor} has its image in the subcategory $\Tele$.
\end{proposition}
\begin{proof}
We must show that for a symmetric monoidal functor $F : \C \to \D$, the induced functor $F : \Optic_\C \to \Optic_\D$ is teleological.

First, $F$ preserves the dualisable morphisms. If $\iota(f, g) : (S, S') \hto (A, A')$,
\begin{align*}
F\iota(f, g) 
&= F(\langle \lambda_A^{-1} f, g \lambda_{A'} \rangle) \\
&= \langle \phi^{-1}_{I,A} (F(\lambda_A^{-1} f)), (F(g \lambda_{A'})) \phi_{I,A'}\rangle \\
&= \langle \phi^{-1}_{I,A} (F\lambda_A^{-1}) (Ff), (Fg)(F \lambda_{A'}) \phi_{I,A'}\rangle \\
&= \langle (\phi^{-1}_I \otimes FA) \phi^{-1}_{I,A} (F\lambda_A^{-1}) (Ff), (Fg)(F \lambda_{A'}) \phi_{I,A'} (\phi_I \otimes FA) \rangle \\
&= \langle \lambda_{FA}^{-1} (Ff), (Fg)\lambda_{FA'} \rangle \\
&= \iota(Ff, Fg)
\end{align*}

And it also preserves the counits:
\begin{align*}
F(\varepsilon_{(S, S')})
&= F(c_{S \otimes S'}) \\
&= F(\langle \rho_S^{-1}, \rho_{S'} \rangle) \\
&= \langle \phi^{-1}_{S,I} (F \rho_S^{-1}), (F \rho_{S'}) \phi_{S',I}\rangle \\
&= \langle (FS \otimes \phi_I^{-1}) \phi^{-1}_{S,I} (F \rho_S^{-1}), (F \rho_{S'}) \phi_{S',I} (FS \otimes \phi_I)\rangle \\
&= \langle \rho_{FS}^{-1}, \rho_{FS'} \rangle \\
&= \varepsilon_{(FS, FS')}
\end{align*}

\todo{This works fine when $\C$ is not strict}
\end{proof}

@@

\begin{proposition}
\label{prop-optic-decompose}
Suppose $\langle l, r \rangle : (S, S') \hto (A, A')$ has complement $M$. Then 
\begin{align*}
\langle l, r \rangle = ((A, I) \switched \varepsilon_{(M, I)} \switched (I, A'))(j(s_{M,A}l) \switched j(rs_{A',M})^*)
\end{align*}
where $j : \C \to \Optic_\C$ is the functor $j(A) := \iota(A, I)$.
\todo{The symmetries could have been avoided in the above expression if the Optic category had been defined with the tensor in the opposite order!}
\end{proposition}
\begin{proof}
First note that because $\C$ is strict monoidal, the counit $\varepsilon_{(M, I)} : (M \otimes I, M \otimes I) = (M, M) \hto (I, I)$ is equal to the connector $c_M : (M, M) \hto (I, I)$.

Then, up to strictness of the monoidal unit, we are composing the two optics
\begin{center}
\input{diagrams/optic-decomposed-outer.tikz}
\end{center}
and
\begin{center}
\input{diagrams/optic-decomposed-inner.tikz}
\end{center}
so the two pairs of twists cancel, and we are left exactly with the diagram for $\langle l, r \rangle$.
\end{proof}

\begin{theorem}
\label{optic-is-free-teleological-cat}
Suppose $\C$ is a strict symmetric monoidal category and $\T$ is a teleological category. For all symmetric monoidal functors $F : \C \to \T_d$, there exists a teleological functor $K : \Optic_\C \to \T$, unique up to isomorphism, such that $Kj \cong F$.
%$\Optic$ is left adjoint to $(-)_d$, so $\Optic_\C$ is the free teleological category on a category of dualisable morphisms $\C$.
\end{theorem}
\begin{proof}
On objects, define $K(S, S') = FS \otimes (FS')^*$. Suppose $\langle l, r \rangle : (S, S') \hto (A, A')$ is an optic. By the previous Proposition,
\begin{align*}
\langle l, r \rangle = ((A, I) \switched \varepsilon_{(M, I)} \switched (I, A'))(j(s_{M,A}l) \switched j(rs_{M,A'})^*)
\end{align*}
But now, in order for $K$ to be a teleological functor with $Kj \cong F$, each piece of this expression must be preserved by $K$, so we are forced to define:
\begin{align*}
K\langle l, r \rangle = (FA \otimes \varepsilon_{FM} \otimes (FA')^*)(F(s_{M,A}l) \otimes (F(rs_{A', M}))^* )
\end{align*}

We must show that this is well defined. Suppose we have two lenses related by the coend relation:
\begin{align*}
\langle (f \otimes A) l, r \rangle = \langle l, r (f \otimes A') \rangle
\end{align*}
Then: \todo{do as string diagram}
\begin{align*}
K\langle (f \otimes A) l, r \rangle 
&= (FA \otimes \varepsilon_{FM} \otimes (FA')^*)(F(s_{M,A}(f \otimes A)l) \otimes (F(rs_{A',M}))^* ) \\
&= (FA \otimes \varepsilon_{FM} \otimes (FA')^*)(F((A \otimes f)s_{N,A}l) \otimes (F(rs_{A',M}))^* ) \\
%&= (FA \otimes \varepsilon_{FM} \otimes (FA')^*)(FA \otimes Ff \otimes FM \otimes FA')(F(s_{N,A}l) \otimes (F(rs_{A',M}))^* ) \\
&= (FA \otimes (\varepsilon_{FM} (Ff \otimes FM)) \otimes (FA')^*)(F(s_{N,A}l) \otimes (F(rs_{A',M}))^* ) \\
&= (FA \otimes (\varepsilon_{FN} (FN \otimes (Ff)^*)) \otimes (FA')^*)(F(s_{N,A}l) \otimes (F(rs_{A',M}))^* ) \\
&= (FA \otimes \varepsilon_{FN} \otimes (FA')^*)(F(s_{N,A}l) \otimes (F(rs_{A',M}(A' \otimes f)))^* ) \\
&= (FA \otimes \varepsilon_{FN} \otimes (FA')^*)(F(s_{N,A}l) \otimes (F(r(f \otimes A')s_{A',N}))^* ) \\
&= K\langle  l, r (f \otimes A') \rangle 
\end{align*}
\todo{functoriality} \todo{monoidal} \todo{preserves dualisable morphisms} \todo{preserves counit}
\end{proof}

@@

Also folklore are \emph{affine traversals}, which allow access to a target that may or may not be present. These are described at \cite{AffineTraversalPost}, and it is hypothesised that they correspond to pairs of maps $\C(S, A \sqcup S') \times \C(S\times A', S')$. This is not precisely correct, but very close:

Suppose $\C$ is cartesian closed, and has binary coproducts. Then let $\mathsf{Aff}$ be the category $\C \times \C$, equipped with the monoidal structure
\begin{align*}
(P', Q') \otimes (P, Q) &= (P' \sqcup (Q' \times P) , Q' \times Q)
\end{align*}
\todo{check $\mathsf{Aff}$ is actually a monoidal category...}

Define a monoidal action $\mathsf{Aff} \to [\C, \C]$ where $(P, Q)$ acts on $A$ by
\begin{align*}
(P, Q) \cdot A = P \sqcup (Q \times A)
\end{align*}
so that as expected,
\begin{align*}
(P', Q') \cdot (P, Q) \cdot A 
&= (P', Q') \cdot  (P \sqcup (Q \times A)) \\
&= P' \sqcup (Q' \times (P \sqcup (Q \times A))) \\
&\cong P' \sqcup (Q' \times P) \sqcup (Q' \times Q \times A) \\
&= (P' \sqcup (Q' \times P) , Q' \times Q) \cdot A \\
&= ((P', Q') \otimes (P, Q)) \cdot A
\end{align*}

\todo{This feels similar to taking some sort of `compositum' of the two actions $\times$ and $\sqcup$, both embed in this category. Marco suggests taking the pushout of the projections into the pullback, calculated in the 2-cat of monoidal categories.}

Now the set of optics $(S, S') \hto (A, A')$ is:
\begin{align*}
\Optic_{\mathsf{Aff}}((S, S'), (A, A')) 
&= \int^{M \in \mathsf{Aff}} \C(S, M \cdot A) \times \C(M \cdot A', S') \\
&\cong \int^{P,Q \in \C} \C(S, (P,Q) \cdot A) \times \C((P,Q) \cdot A', S') \\
&= \int^{P,Q \in \C} \C(S, P \sqcup (Q \times A)) \times \C(P \sqcup (Q \times A'), S') \\
&\cong \int^{P,Q \in \C} \C(S, P \sqcup (Q \times A)) \times \C(P,S') \times \C(Q \times A', S') \\
&\cong \int^{Q \in \C} \C(S, S' \sqcup (Q \times A)) \times \C(Q \times A', S') \\
&\cong \int^{Q \in \C} \C(S, S' \sqcup (Q \times A)) \times \C(Q, \homC(A', S')) \\
&\cong \C(S, S' \sqcup (\homC(A', S') \times A))
\end{align*}

This set includes into $\C(S, A \sqcup S') \times \C(S\times A', S')$, but the two are not equal---in other words---there are fewer affine traversals than expected.

\todo{Give an example of two elements of the set that correspond to the same affine traversal}

\todo{For \emph{lawful} affine traversals, there may not be a difference}

\begin{remark}
For this argument to work, it is important that the morphisms in $\mathsf{Aff}$ are just those in $\C \times \C$, even though there are likely to be more natural transformations between the corresponding functors $P \sqcup (Q \times -)$ and $P' \sqcup (Q' \times -)$.
\end{remark}

@@

So in the above example, we have the product $\times$ acting pointwise on the functor category $\Set \to \Set$. Our example \texttt{first} is then an optic $F \hto G$, where $F = (-) \times \mathtt{Int}$ and $G$ is the identity functor.

@@

\subsection{Indexed Optics}
\todo{
This possibly looks like
\begin{align*}
\int^{M \in \M} \C(S, M(IA)) \times \C(M A', S')
\end{align*}
with another monoidal action $I$. The $I$ would need to distribute over $M$ in some way. ``Coindexed'' optics seem to have 
\begin{align*}
\int^{M \in \M} \C(S, J(MA)) \times \C(M A', S')
\end{align*}
which is why they pass information `in the other direction' but still transform covariantly.
}

@@

@@

\begin{example}
A minimal example of a lens that obeys $\fget\fput$ and $\fput\fget$ but not $\fput\fput$ is the following. Let $p : \{A,B,C\} \hto \{X, Y\}$ be the lens with
\begin{align*}
\fget : \{A,B,C\} &\to \{X, Y\} \\
A &\mapsto X \\
B &\mapsto X \\
C &\mapsto Y 
\end{align*}
\begin{align*}
\fput : \{A,B,C\} \times \{X, Y\} &\to \{A,B,C\} \\
(A,X) &\mapsto A \\
(B,X) &\mapsto B \\
(C,X) &\mapsto B \\
(A,Y) &\mapsto C \\
(B,Y) &\mapsto C \\
(C,Y) &\mapsto C
\end{align*}
\end{example}
